---
title: "In-Class Exercise 02"
description: |
  This is a dummy blog post for testing out the various ins and outs of a Distill blog. In this in-class exercise, I learned how to handle geospatial data in R by using the sf package.
author:
  - name: Megan Sim
    url: https://www.linkedin.com/in/megan-sim-tze-yen/
date: 08-23-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
Getting Started

This code chunk performs three tasks:

- A packaging list call `packages` will be created.  It consists of all the R packages required to accomplish this hands-on exercise.
- Next, the code chunk will check if the R packages in `packages` have been installed in R. If they have yet been installed, they will be installed.
- After all the R packages have been installed, they will be launched in RStudio environment.

``` {r echo=TRUE, eval=TRUE}
packages <- c('sf', 'tidyverse')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

